---
title: "Geography 176A"
author: "[AineRoonan](https://aineroonan.github.io)"
subtitle: 'Lab 06: Flood Risk in Mission Creek: Past, Present, Future'
output:
  html_document:
    theme: cerulean
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	fig.align = "center",
	message = FALSE,
	warning = FALSE,
	out.width = "75%"
)
```

```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
library(tidyverse)
library(sf)       
library(raster)   
library(fasterize) 
library(whitebox)  
library(gifski)
library(knitr)
library(osmdata)   
library(elevatr) 
```

Collecting Data
```{r echo=FALSE, message=FALSE, warning=FALSE, paged.print=FALSE}
basinboundary = read_sf("https://labs.waterdata.usgs.gov/api/nldi/linked-data/nwissite/USGS-11119750/basin")

elevationbasin = elevatr::get_elev_raster(basinboundary, z = 13) %>% 
  crop(basinboundary) %>% 
  mask(basinboundary)

elevationbasin = elevationbasin * 3.281

writeRaster(elevationbasin, "~/github/geog-176A-labs/data/elevationbasin.tif", overwrite = T)

elevation_raster = raster("~/github/geog-176A-labs/data/elevationbasin.tif")

bb_basin = st_bbox(basinboundary) %>% 
  st_as_sfc() %>% 
  st_transform(4326)

buildings = osmdata::opq(bb_basin) %>% 
  add_osm_feature(key = "building") %>% 
  osmdata_sf()

buildings_point = buildings$osm_polygons %>% 
  st_intersection(basinboundary) %>% 
  st_transform(crs(basinboundary)) %>% 
  st_centroid()

buildings_poly = buildings$osm_polygons %>% 
  st_intersection(basinboundary) %>% 
  st_transform(crs(basinboundary))

railstations = opq(bb_basin) %>% 
  add_osm_feature(key = "railway", value = "station") %>% 
  osmdata_sf()

railways = railstations$osm_points %>% 
  st_intersection(basinboundary) %>% 
  st_transform(crs(basinboundary))

osm.bns = opq(bb_basin) %>% 
  add_osm_feature(key = "waterway", value = "stream") %>% 
  osmdata_sf()

osm.stream = osm.bns$osm_lines %>% 
  st_intersection(basinboundary) %>% 
  st_transform(crs(basinboundary))

```

Terrain Analysis
```{r}
wbt_hillshade('~/github/goeg-176A-labs/basin_elev.tif', '~/github/MCH.tif')

```

```{r}
MCH = raster("~/github/MCH.tif")
```


```{r, warning=F, echo=F, message=F}
plot(MCH, col = gray.colors(256, alpha = .5), main = "Hillshade Basin and Stream", legend = F, box = F) 

plot(osm.stream$geometry, add = T,  col = "blue")

plot(basinboundary$geometry, add = T)

osm.stream_buffer = osm.stream %>% 
  st_transform(5070) %>% 
  st_buffer(10) %>% 
  st_transform(4326)

osm.stream_raster = fasterize::fasterize(osm.stream_buffer, elevationbasin)

writeRaster(osm.stream_raster, "~/github/osm.stream-raster.tif", overwrite = T)
```


```{r, eval = F}
wbt_breach_depressions("~/github/geog-176A-labs/data/elevationbasin.tif", "~/github/geog-176A-labs/data/breach_depressions.tif")
```


```{r}
wbt_elevation_above_stream("~/github/breach_depressions.tif", "~/github/osm.stream-raster.tif", "~/github/hand-raster.tif")

```
```{r}
#HAND = raster("~/github/hand-raster.tif")
```
Hello, I am incredibly sorry, but my last line of code would not run on my computer, I have been working on this one issue for well over an hour. I was having the same issues with reading in rasters earlier, but for some reason, those started working, but I have ran into the exact same issues here and I cannot figure out how to fix this issue. I will not have reliable access, if any, to wifi for the rest of the day and week, so this is where I have to end my lab. I am sorry for the unprofessional response. 

