---
title: "Geography 176A"
author: "[AineRoonan](https://aineroonan.github.io)"
subtitle: 'Lab 02: COVID-19 Pandemic'
output:
  html_document:
    theme: journal
---

```{r}
library(tidyverse)
library(knitr)
library(readxl)
library(zoo)
library(kableExtra)
library(dplyr)
library(ggplot2)

url = 'https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv'
covid = read_csv(url)

dat = covid %>% 
  filter(state == "California") %>% 
  group_by(county) %>% 
  mutate(newCases = cases - lag(cases)) %>% 
  ungroup() %>% 
  filter(date == max(date)) 


most_new_cases = dat %>% 
  slice_max(newCases, n = 5) %>% 
  select(county, newCases)

most_cases = dat %>% 
  slice_max(cases, n=5) %>% 
  select(county,cases) 


knitr::kable(most_new_cases,              
        caption = "Most New Cases within California by County",
        col.names = c("County", "New Cases"),
        format.args = list(big.mark = ",")) %>% 
    kableExtra::kable_styling("striped", full_width = TRUE, font_size = 14)

knitr::kable(most_cases,              
      caption = "Most Cases within California by County",
      col.names = c("County", "New Cases"),
      format.args = list(big.mark = ",")) %>% 
    kableExtra::kable_styling("striped", full_width = TRUE, font_size = 14)
```

```{r}

library(readxl)
pop <- read_excel("../data/PopulationEstimates.xls", 
    skip = 2)

popo = 
  pop %>% 
  select(fips = "FIPStxt", state = "State", Area_Name, pop2019 = "POP_ESTIMATE_2019") %>% 
  right_join(covid, by = c("fips"))

cali = popo %>% 
  filter(state.x == "CA") %>% 
  group_by(county) %>% 
  mutate(newCases = cases - lag(cases)) %>% 
  mutate(cpc = cases/pop2019) %>% 
  mutate(cpcnew = newCases/pop2019) %>% 
  ungroup() %>% 
  filter(date == max(date)) 
 
tot = cali %>% 
  group_by(date = max(date)) 
  
  totalcases = sum(tot$cases)
  totalnewCases = sum(cali$newCases)

mostNewCasesbypop = cali %>% 
  arrange((cpcnew)) %>% 
  slice_max(cpcnew, n = 5) %>% 
  select(county, New_Cases_Per_Capita = cpc)

mostCasesbypop = cali %>% 
  slice_max(cpc, n=5) %>% 
  arrange(-cpc) %>% 
  select(county,Cases_Per_Capita = cpc) 

knitr::kable(mostNewCasesbypop,              
        caption = "Most New Cases Per Capita within California by County",
        col.names = c("County", "New Cases Per Capita"),
        format.args = list(big.mark = ",")) %>% 
    kableExtra::kable_styling("striped", full_width = TRUE, font_size = 14)

knitr::kable(mostCasesbypop,              
      caption = "Most Cases per Capita within California by County",
      col.names = c("County", "Cases per Capita"),
      format.args = list(big.mark = ",")) %>% 
    kableExtra::kable_styling("striped", full_width = TRUE, font_size = 14)

fourteen = popo %>% 
  filter(date > max(date) - 14, state.x == "CA") %>% 
  group_by(fips) %>%
  mutate(newCases = cases - lag(cases)) %>% 
  summarize(cases14 = sum(cases, na.rm = TRUE),
            newCases14 = sum(newCases, na.rm = TRUE),
            pop2019 = max(pop2019)) %>% 
  mutate(pc_nc = newCases14 / (pop2019 / 100000)) %>% 
  filter(pc_nc <= 100) %>% 
  select(fips, pc_nc)


# (cases / unit) >= case / unit2
# 100,000 unit1 = unit2

```


#### When I first ran these numbers on Sunday, there were 591,216 new cases cases within California in the last 14 days and 621981 total cases.There were 14 counties that could be considered safe, as they have had less than 100 new cases per 100,000 residents over the past 14 days, but I can see that these numbers have changed, as there are only 13 safe counties now, as is represented by the length of safe$county. Total number of cases is represented by the "totalcases" value and "totalnewCases" represents how many new cases there were within California in the last 14 days. 

```{r}
#Question 2
pop2 = pop %>% 
  select(state= Area_Name, POP_ESTIMATE_2019)
  

four = covid %>% 
  filter(state == "California" | state == "New York" | state == "Louisiana" | state == "Florida") %>%
  group_by(state, date) %>% 
  summarize(cases = sum(cases)) %>% 
  left_join(pop2, by = "state")


morefour = four %>% 
  group_by(state) %>% 
  mutate(newCases = cases - lag(cases)) %>% 
  mutate(avg = rollmean(newCases, 7, fill = NA, align = "right"))

graph1 =
  morefour %>%
  group_by(date) %>% 
  ggplot(aes(x=date, y= newCases, color = state))+
  geom_line(size = 0.5) +
  facet_wrap(~state) +
  ggthemes::theme_fivethirtyeight() +
  labs(title = "Number of Daily New Cases by State",
       x = "Date",
       y = "New Cases",
       caption = "Lab 02")
  ggsave(graph1, file = "../img/graph1.png")

graph2 =
  morefour %>%
  group_by(date) %>% 
  ggplot(aes(x=date, y= avg, color = state))+
  geom_line(size = 0.5) +
  facet_wrap(~state) +
  ggthemes::theme_fivethirtyeight() +
  labs(title = "Average Weekly Increase in Cases by State",
       x = "Date",
       y = "Average Weekly Increases",
       caption = "Lab 02")
  ggsave(graph2, file = "../img/graph2.png")

fourcapita = morefour %>% 
  group_by(state) %>% 
  mutate(newCpTP = newCases/POP_ESTIMATE_2019) %>% 
  mutate(meancap = rollmean(newCpTP, 7, fill = NA, align = "right"))
  
graph3 = fourcapita %>% 
  group_by(date) %>% 
  ggplot(aes(x=date, y= newCpTP, color = state))+
  geom_line(size = 0.5) +
  facet_wrap(~state) +
  ggthemes::theme_economist() +
  labs(title = "Number of Daily New Cases per Capita by State",
       x = "Date",
       y = "New Cases per Capita",
       caption = "Lab 02")
  ggsave(graph3, file = "../img/graph3.png")
 
graph4 = fourcapita %>%
  group_by(date) %>% 
  ggplot(aes(x=date, y= meancap, color = state))+
  geom_line(size = 0.5) +
  facet_wrap(~state) +
  ggthemes::theme_fivethirtyeight() +
  labs(title = "Average Weekly Increase in Cases per Capita by State",
       x = "Date",
       y = "Average Weekly Increases per Capita",
       caption = "Lab 02")
ggsave(graph4, file = "../img/graph4.png")
```

# I think that scaling by population has a large effect on the analysis as it allows us to look at the data in a more meaningful way. If we were to simply compare states by number of cases, it does not take into account the dangers presented as a whole population as it is limited by arbitrary state lines that do not regard density or population. I purposefully set my y to not be free for ease of comparison and for a visual representation. Louisiana looks significantly worse when regarded per capita, as they have a lower population, so statistically, they are less likely to have as many cases as another state, such as New York or California. On the other hand, California looks better when regarded per capita, as they have a larger population, so they were statistically speaking more likely to have more cases.  

```{r}
#Question 3
library(readr)
datas <- read_csv("datas.csv")
View(datas)


```

